### 2004-05-04 Documentation/part2.tex
missing files and plots added

We tell \GLOBES\ that we want to use the built-in neutrino factory
flux, by setting the beam type variable to $1$. The target mass 
is $50\,\mathrm{kt}$ and a total of $ 4.264\cdot10^{21}$
muons is stored. The data is divided into 20 energy bins ranging from
$4\,\mathrm{GeV}$ to $50\,\mathrm{GeV}$. 
\begin{quote}
{\tt /* beam */}\\
{\tt \$beam\_type = 1}\\
{\tt \$target\_mass = 50}\\
{\tt \$stored\_muons = 4.264e+21}\\
{\tt \$numofbins = 20}\\
{\tt \$emin = 4.0}\\
{\tt \$emax = 50.0}
\end{quote}

@@

\GLOBES\ handles only point sources, thus it is not possible
to study effects from the finite size of the neutrino production
region like in the Sun. Furthermore only one source for each experiment
is possible, this make it impossible\footnote{at least rather tricky} 
to simulate multiple source experiments like KamLAND. All neutrino
sources usable within \GLOBES\ therefore can be specified by the
flux spectra for each flavour and CP sign and the total luminosity
of the source. There are two principal ways to provide a flux:
use a built-in source or a file provided by the user.

The built-in source is muon decay, since the neutrino
spectra and flavour composition are known analytically in that case. For 
this sources the energy of the parent particle has to be specified
and the number of useful decays per year.

In the latter case the fluxes are read from a file with seven rows and
501 lines.
\begin{quotation}
$ E\quad
\Phi_{\nu_e}\quad
\Phi_{\nu_\mu}\quad
\Phi_{\nu_\tau}\quad
\Phi_{\bar\nu_e}\quad
\Phi_{\bar\nu_\mu}\quad
\Phi_{\bar\nu_\tau}$
\end{quotation}
where the energy values are evenly spaced. 
For accessing fluxes at 
arbitrary energies linear interpolation is used. Once the energy leaves the
range of values given in the file $0.0$ is returned. The units used for the
fluxes are not specified and right now each flux has its own fudge factor
which is hard encoded in the file {\tt nufact.c}. The problem is that
in computing the normalization one has to know things like number of target
particles per unit mass asf. which is not always straightforward. Thus a fudge
factor will be useful in any case. Right now however this is not the only 
fudge factor since for some experiments the definition of their target mass
involves another fudge factor.

Since many sources have two modes of operation, it is possible to define
a pair of fluxes by using the polarity flag when loading the files. Usually
files connected in this way use the suffixes ``plus'' and ``minus'' in the 
file names. The individual members of the pair can be accessed by the polarity
variable. Also the built-in fluxes provide this choice.

For sources based on $\pi$-decay and nuclear reactors the total power
involved defines how many neutrinos are produced therefore this variable is
used to define the luminosity. Furthermore it is useful to define 
the operation time of an experiment in years. Note however that the 
number of events used in analysis only depends on the product of
\begin{equation}
\mathrm{mass}\,\left[10^6\,\mathrm{kg}\right]\times \mathrm{time}
\,\left[\mathrm{years}\right]\times\left\{ \begin{array}{c}
\mathrm{source~power}\,\left[\mathrm{MWy}\right]\\
\mathrm{useful~decays}\,\left[\mathrm{year}^{-1}\right]
\end{array}\right.\,.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Oscillation properties}

\subsection{Baseline \& matter profile}
The purpose of this module is to specify the path the neutrinos
are traveling. The central piece of information is the total length
of the path - the baseline\index{Baseline}
\begin{equation}
L\,\left[\mathrm{km}\right]
\end{equation}

Furthermore there is
the matter profile encountered along the baseline which can be set in
several ways. The simplest way is to use a constant density corresponding
to the average density along the baseline calculated using the 
PREM\cite{Stacey} onion shell model of the 
Earth\index{PREM}\index{Earth matter density}.

Another possibility is to use several layers of constant density each in 
order to improve the approximation to the actual Earth matter profile. In this
case the user specifies the total baseline and the number of layers. The
time consumed in computing the oscillation probabilities is directly 
proportional to the number of layers.

The third possibility to specify the matter profile is to provide a list
of thicknesses and densities for the layers.

@@

The most obvious property of the neutrino detection is the total
active mass of the target as already defined in section~\ref{sec:source}.

@@

@@

!!!!!!!!!!!!
Do we need $\epsilon(E')$ ???
!!!!!!!!!!!!

@@

\subsection{Rules}
\label{sec:rules}

@@

Before going into the details we define that the range from $E_\mathrm{min}$
to $E_\mathrm{max}$ is divided in a certain number of bins $N_\mathrm{bins}$.
These bins form a partition, which not necessarily is equidistant. The 
partition is any case completely defined by specifying $E_\mathrm{min}$
and $E_\mathrm{max}$ and the size of each bin  $\Delta E_i$. It is also
straightforward to compute the midpoint of each bin $E_i$.
In order to obtain the number of events 
$n_i^c$ from one channel $c$ in one bin $i$ one has to solve this integral
%
\begin{eqnarray}
\label{eq:events_bin}
n_i^c=N/L^2\,\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad \int dE \,\, \Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
R^c(E,E')\,
\epsilon^c(E')\,.
\end{eqnarray} 
%
The integration with respect to the reconstructed energy $E'$ can be
performed independently of the oscillation parameters and we define
the bin kernel for the $i$th bin $K_i^c$
\begin{equation}
\label{eq:kernel}
K_i^c(E):=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad R^c(E,E')\,
\epsilon^c(E')\,.
\end{equation}
With this definition equation~\ref{eq:events_bin} can be rewritten as
\begin{equation}
\label{eq:simple_int}
n_i^c=N/L^2 \int dE\quad  \underbrace{\Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
K_i^c(E)\,}_{f(E)}. 
\end{equation}

>>>>

We tell \GLOBES\ that we want to use the built-in neutrino factory
flux, by setting the beam type variable to $1$. The target mass 
is $50\,\mathrm{kt}$ and a total of $ 4.264\cdot10^{21}$
muons is stored. The data is divided into 20 energy bins ranging from
$4\,\mathrm{GeV}$ to $50\,\mathrm{GeV}$. 
\begin{quote}
{\tt /* beam */}\\
{\tt \$beam\_type = 1}\\
{\tt \$target\_mass = 50}\\
{\tt \$time = 8}\\
{\tt \$stored\_muons = 5.33e+20}\\
{\tt \$numofbins = 20}\\
{\tt \$emin = 4.0}\\
{\tt \$emax = 50.0}
\end{quote}

@@

\GLOBES\ handles only point sources, thus it is not possible
to study effects from the finite size of the neutrino production
region like in the Sun. Furthermore only one source for each experiment
is possible, this make it impossible\footnote{at least rather tricky} 
to simulate multiple source experiments like KamLAND. All neutrino
sources usable within \GLOBES\ therefore can be specified by the
flux spectra for each flavour and CP sign and the total luminosity
of the source. There are two principal ways to provide a flux:
use a built-in source or a file provided by the user. Built-in sources
as well as user defined fluxes are referred to internally by the
 {\tt \$beam\_type} variable. Certain values of  {\tt \$beam\_type} are
pre-defined and used for the built-in fluxes, \eg\ {\tt \$beam\_type = 1}
selects the neutrino factory flux.

The built-in source is muon decay, since the neutrino
spectra and flavour composition are known analytically in that case. For 
this sources the energy of the parent particle has to be specified
and the number of useful decays per year. Thus the sequence for setting
up a neutrino factory is
\begin{quote}
{\tt \$beam\_type = 1}\\
{\tt \$parent\_energy = 50.0}\\
{\tt \$stored\_muons = 5.33e+20}\\
\end{quote}
The values of {\tt \$parent\_energy} and {\tt emax} are always the same.

User defined fluxes are dealt with a little different at initialization
and the value of {\tt \$beam\_type} is automatically assigned. This is
necessary to avoid any problems with multiple experiments, since the fluxes
are read only once. Reading a flux file looks like
\begin{quote}
{\tt flux(\#dummy)<}\\
{\tt \hspace*{3ex}@flux\_file\_p = user\_file\_1.dat}\\
{\tt \hspace*{3ex}@flux\_file\_m = user\_file\_2.dat}\\
{\tt >}\\
{\tt \$target\_power = 4.0 }
\end{quote}

Since many sources have two modes of operation, it is possible to define
a pair of fluxes by using the polarity flag when loading the files. Usually
files connected in this way use the suffixes ``plus'' and ``minus'' in the 
file names. The individual members of the pair can be accessed by the polarity
variable. Also the built-in fluxes provide this choice.
The flux read by {\tt @flux\_file\_p} can be accesed by selecting positive
polarity, whereas the flux read by {\tt @flux\_file\_m} is selected by a 
negative value of the polarity.

For sources based on $\pi$-decay and nuclear reactors the total power
involved defines how many neutrinos are produced therefore this variable is
used to define the luminosity. Furthermore it is useful to define 
the operation time of an experiment in years. Note however that the 
number of events used in analysis only depends on the product of
\begin{equation}
\mathrm{mass}\,\left[10^6\,\mathrm{kg}\right]\times \mathrm{time}
\,\left[\mathrm{years}\right]\times\left\{ \begin{array}{c}
\mathrm{source~power}\,\left[\mathrm{MWy}\right]\\
\mathrm{useful~decays}\,\left[\mathrm{year}^{-1}\right]
\end{array}\right.\,.
\end{equation}

The format of the flux files is seven rows and
501 lines.
\begin{quotation}
$ E\quad
\Phi_{\nu_e}\quad
\Phi_{\nu_\mu}\quad
\Phi_{\nu_\tau}\quad
\Phi_{\bar\nu_e}\quad
\Phi_{\bar\nu_\mu}\quad
\Phi_{\bar\nu_\tau}$
\end{quotation}
where the energy values are evenly spaced. 
For accessing fluxes at 
arbitrary energies linear interpolation is used. Once the energy leaves the
range of values given in the file $0.0$ is returned. The units used for the
fluxes are not specified and right now each flux has its own fudge factor
which is hard encoded in the file {\tt nufact.c}. The problem is that
in computing the normalization one has to know things like number of target
particles per unit mass asf. which is not always straightforward. Thus a fudge
factor will be useful in any case. Right now however this is not the only 
fudge factor since for some experiments the definition of their target mass
involves another fudge factor. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Oscillation properties}

\subsection{Baseline \& matter profile}
The purpose of this module is to specify the path the neutrinos
are traveling. The central piece of information is the total length
of the path - the baseline\index{Baseline}
\begin{equation}
L\,\left[\mathrm{km}\right]
\end{equation} 
The baseline is set by 
\begin{quote}
{\tt \$baseline = 3000.0 }
\end{quote}

Furthermore there is
the matter profile encountered along the baseline which can be set in
several ways. The simplest way is to use a constant density corresponding
to the average density along the baseline calculated using the 
PREM\cite{Stacey} onion shell model of the 
Earth\index{PREM}\index{Earth matter density}.
\begin{quote}
TBD
\end{quote}

Another possibility is to use several layers of constant density each in 
order to improve the approximation to the actual Earth matter profile. In this
case the user specifies the total baseline and the number of layers. The
time consumed in computing the oscillation probabilities is directly 
proportional to the number of layers.
\begin{quote}
TBD
\end{quote}

The third possibility to specify the matter profile is to provide a list
of thicknesses and densities for the layers. 
\begin{quote}
{\tt \$densitytab=\{ 2.8, 3.5\}}\\
{\tt \$lengthtab=\{ 1000.0, 2000.0\}}\\
\end{quote}
It is important that both list are of equal length and that the sum
of the thicknesses given in  {\tt \$lengthtab} add up to the value of
{\tt \$baseline}.

@@

The most obvious property of the neutrino detection is the total
active mass of the target as already defined in section~\ref{sec:source} and
its value is defined by 
\begin{quote}
{\tt \$target\_mass = 50.0 }
\end{quote}
Furthermore the running time has to be specified by
\begin{quote}
{\tt \$time = 8.0 }
\end{quote}

@@

A cross section file is read by 
\begin{quote}
{\tt flux(\#name)<}\\
{\tt \hspace*{3ex}@cross\_file =user\_file\_1.dat}\\
{\tt >}
\end{quote}  
and can later be refered to by {\tt \#name}.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Rules}
\label{sec:rules}

@@

Before going into the details we define that the range from $E_\mathrm{min}$
to $E_\mathrm{max}$ is divided in a certain number of bins $N_\mathrm{bins}$.
These bins form a partition, which not necessarily is equidistant. The 
partition is any case completely defined by specifying $E_\mathrm{min}$
and $E_\mathrm{max}$ and the size of each bin  $\Delta E_i$. It is also
straightforward to compute the midpoint of each bin $E_i$.
In order to obtain the number of events 
$n_i^c$ from one channel $c$ in one bin $i$ one has to solve this integral
%
\begin{eqnarray}
\label{eq:events_bin}
n_i^c=N/L^2\,\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad \int dE \,\, \Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
R^c(E,E')\,
\epsilon^c(E')\,.
\end{eqnarray} 
%

Instead of using directly equation~\ref{eq:e_res} we will use a different
definition concerning $\epsilon(E')$. In general  $\epsilon(E')$ has to be
determined by means of a Monte Carlo simulation of the experiment. This
usually involves anyhow a binning of the simulated events in the reconstructed
energy $E'$, therefore one more easily can obtain  $\hat\epsilon_i$ defined
by
\begin{equation}
\label{eq:post_smearing}
\hat\epsilon_i \cdot \int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad R^c(E,E')\,.
\end{equation}
These  $\hat\epsilon_i$ are refered to as so called 
``post-smearing'' efficiencies.

The integration with respect to the reconstructed energy $E'$ can be
performed independently of the oscillation parameters and we define
the bin kernel for the $i$th bin $K_i^c$
\begin{equation}
\label{eq:kernel}
K_i^c(E):=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad R^c(E,E')\,.
\end{equation}
With this definition equation~\ref{eq:events_bin} can be rewritten as
\begin{equation}
\label{eq:simple_int}
n_i^c=N/L^2 \int dE\quad  \underbrace{\Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
K_i^c(E)\,}_{f(E)}. 
\end{equation}
